### 2019-12-19 sections/introduction.tex
fleshing out method and results now anaylsis is complete

% BEGIN HERE
Spiral structure is present in a majority of massive galaxies \citep{2008MNRAS.389.1179L}, yet the formation mechanisms through which spiral structure originates are still hotly debated. Spirals themselves are as diverse as the theories proposed to govern their evolution; ranging from the quintessential pair of well-defined arcs of the grand design spiral, to the patchy and fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral. These variatons on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}). Our current understanding of the mechanisms which drive spiral growth and evolution suggest that each of the different forms of spiral galaxy may be triggered primarily by different processes. Grand Design spirals are thought to have undergone a tidal interaction, be driven by a bar (as suggested by Manifold theory, \citealt{2011hsa6.conf..548R}), or be obeying quasi-stationary density wave theory, in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}. Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and reccurent in nature \citep{1966ApJ...146..810J}. However it is recognised that no two methods of spiral formation are mutually exclusive.

@@

Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (i.e. \citealt{2015A&A...582A..86H}), fourier analysis (i.e. \citealt{2019arXiv190804246D}), more complex automated identification (i.e. SpArcFiRe, \citealt{2014ApJ...790...87D}), and combinations of automated methods and human classifiers \citep{2017MNRAS.472.2263H}. One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} \comment{(Lingard et al, in prep)}. These methods attempt to localize light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization.

This paper makes use of the photometric models obtained through the \textit{Galaxy Builder} citizen science project \comment{(Lingard et al, in prep)} which measure spiral brightness, thickness and tightness. In this paper we focus on the use of measured spiral tightness (quantified using pitch angle, \citealt{1987gady.book.....B}) as a probe into the dynamical mechanisms governing a spiral galaxy's evolution. We make use of multilevel modelling to test for signs of spiral arm winding using the predictions derived by \citealt{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$) and conclude that our sample of galaxies do not show winding of this form. We model galaxy pitch angle as a truncated normal distribution make use of the resulting pitch angles to examine the correlation between pitch angle and bulge size, implied by the Hubble sequence, and bar strength, implied by Manifold theory.

>>>>

% BEGIN HERE
Spiral structure is present in a majority of massive galaxies (\citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}), yet the formation mechanisms through which spiral structure originates are still hotly debated. Spirals themselves are as diverse as the theories proposed to govern their evolution; ranging from the quintessential pair of well-defined arcs of the grand design spiral, to the patchy and fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral. These variatons on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}). Our current understanding of the mechanisms which drive spiral growth and evolution suggest that each of the different forms of spiral galaxy may be triggered primarily by different processes. Grand Design spirals are thought to have undergone a tidal interaction, be driven by a bar (as suggested by Manifold theory, \citealt{2011hsa6.conf..548R}), or be obeying quasi-stationary density wave theory, in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}. Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and reccurent in nature \citep{1966ApJ...146..810J}. However it is recognised that no two methods of spiral formation are mutually exclusive.

@@

\subsection{Measuring galaxy pitch angle}

Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (\citealt{2015A&A...582A..86H}), fourier analysis (i.e. \textsc{2DFFT}, \citealt{2012ApJS..199...33D}), texture analysis (i.e. SpArcFiRe, \citealt{2014ApJ...790...87D}), and combinations of automated methods and human classifiers \citep{2017MNRAS.472.2263H}. One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} (\comment{Lingard et al, in prep}). These methods attempt to localize light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization. This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}. \comment{Lingard et al, in prep} successfully solve this problem through the use of citizen science to provide the starting point for a computational fit.

A common assumption when measuring galaxy pitch angle is that observed spiral arms have a constant pitch angle, these spirals are known as logarithmic spirals and are described by

\begin{equation}
r = A\,e^{\theta\tan\phi}.
\end{equation}

One method used to obtain a pitch angle of a galaxy from identified arm segments is by taking the weighted mean of their pitch angles (as used by \textsc{SpArcFiRe}, \citealt{2014ApJ...790...87D}), where weighting is determined by the length of the arc segment, with longer arcs having higher weights, i.e. for a galaxy where we have identified $N$ arm segments, each with length $L_i$ and pitch angle $\phi_i$

\begin{equation}
  \phi_\mathrm{gal} = \left(\sum_{i=1}^{N}L_i\right)^{-1}\sum_{i=1}^{N}L_i \phi_i.
\end{equation}

The most commonly used measurement of uncertainty of length-weighted pitch angles is the \comment{unweighted?} sample variance between the arm segments identified.

One notable drawback of the use of length-weighted arms is the sensitivity of the end result to the number and quality of the spiral arm segments identified \comment{...}

This paper makes use of the photometric models obtained through the \textit{Galaxy Builder} citizen science project for the 196 spiral galaxies present in \comment{Lingard et al, in prep}. In this paper we focus on the use of measured spiral tightness (quantified using pitch angle, \citealt{1987gady.book.....B}) as a probe into the dynamical mechanisms governing a spiral galaxy's evolution. We make use of Bayesian hierarchical modelling to measure galaxy pitch angle from the spiral arm clusters producted by \textit{Galaxy Builder}. We test for signs of spiral arm winding using the predictions derived by \citealt{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$) and conclude using a marginalized Anderson-Darling test that we \comment{cannot unilaterally reject winding of this form at the 1\% level}. Section \ref{section:morphology_comparision} examines the correlation between pitch angle and bulge size, implied by the Hubble sequence, and bar strength, implied by Manifold theory, and find \comment{no significant correlation}.

### 2019-12-19 sections/method.tex
fleshing out method and results now anaylsis is complete

% !TEX root = ../main.tex

\subsection{Length-weighted spiral arms}

A common method used to obtain a pitch angle of a galaxy is by identifying all

@@

We fit a number of models, including various priors on the global pitch angle distribution. We make use of PYMC3\footnote{\url{https://docs.pymc.io/}}, an open source probabilistic programming framework written in python.

\subsubsection{Spiral arms from \textit{Galaxy Builder}}

Clustered, cleaned points in polar coordinates

$$ r_\mathrm{arm} = \exp\left[\theta\tan(\phi_\mathrm{arm}) + c_\mathrm{arm}\right] + \sigma_r.$$

\subsubsection{Cot-Uniform model}

Assume the pitch angle of a galaxy's spiral arms are drawn from a Normal distribution, truncated between 0 and 90, centred on the galaxy's ``pitch angle'', $\phi_\mathrm{gal}$ with some measure of spread, $\sigma_\mathrm{gal}$, common to all galaxies:

@@

We choose a prior on $\phi_\mathrm{gal}$ such that

\begin{equation}
\cot\left(\phi_\mathrm{gal}\right) \sim \mathrm{Uniform}(\mathrm{min}=1, \mathrm{max}=4),
\end{equation}

and $\sigma_\mathrm{gal}$ of

\begin{equation}
\sigma_\mathrm{gal} \sim \mathrm{InverseGamma}(\alpha=2,\,\beta=20).
\end{equation}

We model point

\begin{equation}
c_\mathrm{arm} \sim \mathrm{Cauchy}(\alpha=0,\,\beta=10).
\end{equation}

\begin{equation}
\sigma_r \sim \mathrm{HalfCauchy}(\beta=0.2)
\end{equation}

\subsubsection{Hierarchial Normal Model}

Instead of the above model, assume

\begin{equation}
\phi_\mathrm{gal} \sim \mathrm{TruncatedNormal}(\mu_\mathrm{global}, \sigma_\mathrm{global}, \mathrm{min}=0, \mathrm{max}=90),
\end{equation}

Hyperpriors used are as follows:

\begin{equation}
\mu_\phi \sim \mathrm{Uniform}(0, 90)
\end{equation}

\begin{equation}
\sigma_\mathrm{global} \sim \mathrm{InverseGamma}(\alpha=1,\,\beta=10)
\end{equation}

>>>>

% !TEX root = ../main.tex
\subsection{The Galaxy Sample}
The galaxies selected analysed in this paper are the 198 galaxies from \comment{Lingard et al, in prep.}. We combine classifications of galaxies which were repeated in the \textit{validation subset} with the original classifications, and perform clustering and point cleaning as detailed in \comment{Lingard et al, in prep.}, and remove any galaxies for which no arms were identified. This results in a hierarchiacal data structure of 109 galaxies and 250 spiral arms, with 307,861 points in polar coordinates, which are scaled such that the radius of each spiral arm has unit maximum.

\comment{Should I elaborate more on the sample?}

@@

Assume we can model spiral arms as a logarithmic spiral, described by

\begin{equation}
r_\mathrm{arm} = \exp\left[\theta_\mathrm{arm}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right].
\end{equation}

Assume the pitch angle of a galaxy's spiral arms are drawn from a Normal distribution, truncated between 0 and 90, centred on some value we will call the galaxy's pitch angle, $\phi_\mathrm{gal}$. This dispersion in arm pitch angle has some measure of spread, $\sigma_\mathrm{gal}$, which we will assume is the same in galaxies:

@@

We choose priors on $\phi_\mathrm{gal}$ and $\sigma_\mathrm{gal}$ of

\begin{align}
  \phi_\mathrm{gal} &\sim \mathrm{Uniform}(\mathrm{min}=0, \mathrm{max}=90),\\
  \sigma_\mathrm{gal} &\sim \mathrm{InverseGamma}(\alpha=2,\,\beta=20).
\end{align}

We also have the offset parameter $c$, and a measure of radial uncertainty $\sigma_r$:

\begin{align}
  c_\mathrm{arm} &\sim \mathrm{Cauchy}(\alpha=0,\,\beta=10),\\
  \sigma_r &\sim \mathrm{HalfCauchy}(\beta=0.2).
\end{align}

\begin{figure}
  \includegraphics[width=8cm]{plots/plots_n109d1000t500/model.pdf}
  \caption{The model used for galaxy pitch angle measurement for the sample.}
  \label{fig:ad-cot-test}
\end{figure}

To perform inference, we make use of the No-U-Turn-Sampler (NUTS), implemented in PYMC3\footnote{\url{https://docs.pymc.io/}}, an open source probabilistic programming framework written in python. \comment{does PYMC3 have a citation?}
